---
title: "Homework 3"
output: github_document
author: Sushupta Vijapur
---
```{r setup}
library(tidyverse)
library(p8105.datasets)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

## all plots i make will have the viridis color palette
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```


## Problem 1

Load "instacart" dataset

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and ... columns 

Observations are the level of itesm in the order by users. There are user / order variables -- user id, order id, order day and order hour. There are also item variables -- name, aisle, department and some numeric codes.

### Part 1
How many aisles and which are the most items from?

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

### Part 2
Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
Counting produces a dataframe that we can manipulate

Notes for me: 
We want to rotate axis labels because they're all overlapping 
Order from least to most number of orders (change aisle to a factor)
  "aisle = fct_reorder(aisle, n)" reorder aisle by n
```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

### Part 3
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

Notes for me:
To do a ranking that is different for each aisle, we need to group first
Putting group_by before count is important because in the output we will retain the aisle. So it will count the n BY the aisle.
```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(aisle, rank)) %>% 
  knitr::kable()
  
```

### Part 4
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```

## Problem 2

### Part 1
Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
accel_data = 
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
   pivot_longer(
    activity_1:activity_1440,
    names_to = "min_of_day",
    names_prefix = "activity_",
    values_to = "activity_count"
  ) %>% 
  mutate(
    activity_count = as.numeric(activity_count),
    min_of_day = as.numeric(min_of_day),
    day_type = ifelse(day == "Saturday" | day == "Sunday", "weekend", "weekday"),
    day = factor(day),
    day = fct_relevel(day, "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
  )

```
The results data frame has `r nrow(accel_data)` rows and `r ncol(accel_data)` columns. The data frame includes week, day_id and day of the week variables. I used the day variable to create a day_type variable which indicates whether it is a weekday or a weekend. Furthermore, I pivoted the activity counts and minutes variables to longer. Now there is an activity count and min of the day variable in tidy format.Furthermore, by re-leveling the data, the days of the week will be listed in intuitive order. Lastly, there is a week variable which numbers the weeks during the observational period (1-5 weeks). All this data was collected from a 63 year old man with a BMI of 25 who was diagnosed with congestive heart failure. 


### Part 2
Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied data-set, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel_data %>% 
  group_by(day, week) %>% 
  summarize(total_act = sum(activity_count)) %>% 
  pivot_wider(
    names_from = day,
    values_from = total_act
  ) %>% 
   knitr::kable()
```
In weeks 4 and 5, the activity was particularly lower than in other weeks and other days. 


### Part 3
Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel_data %>% 
  ggplot(aes(x = min_of_day, y = activity_count, color = day)) +
  geom_line() +
  geom_smooth(aes(group = day)) +
  labs(
    title = "Activity Plot",
    x = "Minutes of the day",
    y = "Activity Count"
  )
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```
The plot above shows minutes of the day and activity counts by days of the week (colors). By including a smooth line to show the average trend by day, we are able to observe patterns more clearly. There seems to be a noticeable peak in activity counts on Friday evening and around the middle of the day on Sunday. Other than those clear peaks in the curves, the rest of the activity counts seem to be around the same throughout the day for each day of the week.



## Problem 3

### Part 1
Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?

Load data

```{r}
data("ny_noaa")
```

Clean data

```{r}
ny_noaa =
  ny_noaa %>% 
  separate(date, c("year", "month", "day"), convert = TRUE) %>% 
  mutate(
    prcp = prcp / 100,
    snow = snow / 10,
    snwd = snwd / 10,
    tmax = as.numeric(tmax) / 10,
    tmin = as.numeric(tmin) / 10
  )
   

month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

tidy_ny_noaa = left_join(ny_noaa, month_df, by = "month")

ny_noaa %>% 
  count(snow) %>% 
  arrange(desc(n))
```
This dataset contains `r nrow(tidy_ny_noaa)` rows and `r ncol(tidy_ny_noaa)` columns. I merged in a month names variable to list the names of the months instead of the numbers. FUrthermore, I converted tmin (minimum temperatue), tmax (maximum temperature) and prcp (precipitation) to acceptable and readable units. Furthermore, I split up the date variable into day, month and year. This dataset also contains data on snowfall and snowdepth as well as the stations across New York State reporting these numbers. 

The most commonly observed value for snowfall in mm is 0. This is probably due to the fact that the data is collected from NYS. For the majority of a given year, there is not snow (Spring, Summer and most of Fall). 


### Part 2
Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?


```{r}
tidy_ny_noaa %>% 
  group_by(id, year, month_name) %>% 
  filter(month_name %in% c("January", "July")) %>% 
  summarize(
    mean_max_temp = mean(tmax, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = year, y = mean_max_temp, group = id)) +
  geom_point(size = 0.1) +
  geom_path(size = 0.1) +
  facet_grid(. ~ month_name) +
  labs(
    title = "Average mean temperature for January and July by Station",
    x = "Year",
    y = "Average Max Temp (c)"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
  
```
Yes, there are two outliers in the July plot around 2005 and 2007, where one is above the plot and the other is below the plow respectively. Most of the mean temps lie between 10 and 30 degrees C. Therefore, there are also some values that are far from the general average mean temp. For example, in the January plot, around 1999, the average mean temp was below -10 degrees C with would be considered an outlier. There are several other instances like this in both months.The stations do overlap over the time period in terms of max and min temps. Lastly, the mean temp in July was higher than the mean temp in January.

### Part 3
Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year
 

```{r}
hex_plot = 
ny_noaa %>% 
  ggplot(aes(x = tmax, y = tmin)) +
  geom_hex() +
  theme(legend.direction = "vertical") +
  labs(
    title = "Max temp versus Min temp",
    x = "Maximum temperature",
    y = "Minimum temperautre"
  )
  

violin_plot = 
  ny_noaa %>% 
  filter(snow < 100, snow > 0, na.rm = TRUE) %>%
  group_by(year) %>% 
  mutate(
    year = as.factor(year)
  ) %>% 
  ggplot(aes(x = year, y = snow)) +
  geom_violin(aes(fill = year), alpha = 0.5) +
  labs(
    title = "Distribution of snowfall (0-100mm) by year",
    x = "Snowfall (mm)",
    y = "Years"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
  
hex_plot + violin_plot 
```
From the hex plot of tmax versus tmin, we can see that while there is some variability, the majority of the values are at the center of the distribution. 

From the violin plot we can see that was not a lot of the variability in snowfall across the years of the study. 
